---
title: "Project"
output: pdf_document
date: "2022-11-29"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(lubridate)
library(dplyr)
library(zoo)
library(readxl)
library(astsa)
library(ggplot2)
```


```{r}
cpi = read.csv("CPI(ALL ITEMS).csv")
cpi.gas = read.csv("cpi(gasoline).csv")
crudeoil = read.csv("crude-oil-price.csv")
gasoline.retail = read.csv("gasoline retail price.csv")
production = read.csv("OPEC_Crude_Oil.csv",header=FALSE)
```

## Data Pre-processing
```{r}
crudeoil$date <- as.Date(crudeoil$date)
```

Arrange from least recent to most recent
```{r}
cpi.gas <- cpi.gas %>% arrange(mdy(cpi.gas$DATE))
crudeoil <- crudeoil[order(crudeoil$date),]
gasoline.retail <- gasoline.retail %>% arrange(my(gasoline.retail$Date))
```

```{r}
cpi.ts = ts(cpi$CPIAUCSL, start=c(2012,1), end=c(2022,9),frequency = 12)
cpi.gas.ts = ts(cpi.gas$CUUR0000SETB01, start=c(2012,1), end=c(2022,9),frequency=12)
crudeoil.ts = ts(crudeoil$price.barrel.42.gal., start=c(2012,1), end=c(2022,9),frequency=12)
gasoline.ts = ts(gasoline.retail$U.S..All.Grades.All.Formulations.Retail.Gasoline.Prices..Dollars.per.Gallon., start=c(2012,1),  end=c(2022,9),frequency=12)
production.ts = ts(production$V2,start=c(2012,1),  end=c(2022,9),frequency=12)
```

```{r}
plot(gasoline.ts, ylab="Price per Gallon ($)", xlab = "Year", main = "Gasoline Monthly Retail Prices")
t = aggregate(gasoline.ts,FUN=mean)
autoplot(t)

Acf(gasoline.ts,50)
Pacf(gasoline.ts)

autoplot(cpi.ts)
autoplot(cpi.gas.ts)
autoplot(crudeoil.ts)
autoplot(production.ts)
```


```{r}
df1 = gasoline.ts 
df2 = cpi.ts 
df3 = cpi.gas.ts
df4 = crudeoil.ts
df5 = production.ts

for (i in 2:5) {
  df1 <- cbind(df1, eval(parse(text=paste("df", i, sep = ''))))
}

colnames(df1) <- c('Gasoline_Retail_Price','CPI(ALL)',
                   'CPI(Gas)',
                   'Crude_Oil_Price',
                   'Crude_Oil_Production')
df1 = as.data.frame(df1)
```

## Simple Naive & Seasonal Naive
```{r}
gasoline.ts <- ts(df1$Gasoline_Retail_Price, start=c(2012,1),end=c(2022,9),frequency = 12)

n=length(gasoline.ts)
stepsAhead = 9
nTraining= n-stepsAhead
training <- window(gasoline.ts, start=c(2012,1), end=c(2012,nTraining))
validation <- window(gasoline.ts, start=c(2012,nTraining+1), end=c(2012,nTraining+stepsAhead))

naive = naive(training, h= stepsAhead)

autoplot(training, main = "Forecasts from Simple Naive Method") + autolayer(validation) + autolayer(naive, PI=FALSE, series = "Predictions")
accuracy(naive, validation)
```

```{r}
snaive = snaive(training, h= stepsAhead)
autoplot(training, main = "Forecasts from Seasonal Naive Method") + autolayer(validation) + autolayer(snaive, PI=FALSE, series = "Predictions")
accuracy(snaive, validation)
```

## Moving Average Model
```{r}
gasoline.ts <- ts(df1$Gasoline_Retail_Price, start=c(2012,1),end=c(2022,9),frequency = 12)

n = length(gasoline.ts)

w = 9
nValid = w

ma.trail.pred= rep(NA, nValid)

for (i in 1:nValid) {
  nTraining= n-nValid + (i - 1)
  training1=window(gasoline.ts, start=c(2012, 1), 
                    end=c(2012, nTraining))
  
  ma.trailing.tmp = rollmean(training1, k=w, align="right")
  last.ma=tail(ma.trailing.tmp,1)
  
  ma.trail.pred[i] = last.ma
}

ma.pred= ts(ma.trail.pred,start=c(2012,length(gasoline.ts)-nValid+1), frequency = 12)

validation <- window(gasoline.ts, start=c(2022,1))

ma.trailing.right = rollmean(training, k=9, align="right")
ma.trailing.center = rollmean(training, k=9, align="center")

autoplot(training, main = "Forecasts from Rolling Moving Average Method") +
  autolayer(validation, series ="Validation")+autolayer(ma.pred, series = "Prediction") +
  autolayer(ma.trailing.right, series = "Trailing MA") +
  autolayer(ma.trailing.center, series = "Centered MA")

e <- ma.pred - validation 

m <- quantile(e, probs = c(0.025,0.975))
lb=ma.pred+m[1]
ub=ma.pred+m[2]

autoplot(ma.pred) +
  autolayer(lb)+ autolayer(ub)

accuracy(ma.trailing.right, training)
accuracy(ma.pred, validation)
```

## Holt-Winter's model
```{r}
gasoline.ts <- ts(df1$Gasoline_Retail_Price, start=c(2012,1),end=c(2022,9),frequency = 12)

n=length(gasoline.ts)
stepsAhead = 9

nTraining=n-stepsAhead
training2 <- window(gasoline.ts, start=c(2012,1), end=c(2012,nTraining))

zzz=ets(training2,model='ZZZ')
zzz.pred=forecast(zzz,h=stepsAhead,level=0)
autoplot(training2, main = "Forecasts from Holt-Winter's Method")+autolayer(validation, series = "Validation")+autolayer(zzz.pred, series = "Predictions")
accuracy(zzz.pred,validation)
```

## ARIMA Model
```{r}
par(mfrow=c(1,2))
Acf(training)
Pacf(training)
par(mfrow=c(1,1))
```

```{r}
auto.arima(gasoline.ts)
```

```{r}
m1 = arima(training,order=c(0,1,1),seasonal=list(order=c(1,0,0),period=12))
m1.pred = forecast(m1,h=9)
accuracy(m1.pred, validation)
```

```{r}
m2 = arima(training,order=c(0,1,1),seasonal=list(order=c(0,0,0),period=12))
m2.pred = forecast(m2,h=9)
accuracy(m2.pred, validation)

autoplot(training, main = "Forecast from ARIMA(0,1,1)") +
  autolayer(validation, series="Validation")+
  autolayer(m2.pred, series="Prediction",PI=FALSE)
```

## Neural Network
```{r}
gasoline.ts <- ts(df1$Gasoline_Retail_Price, start=c(2012,1),end=c(2022,9),frequency = 12)

nValid = 9
nTrain <- length(gasoline.ts) - nValid
train.ts <- window(gasoline.ts, start = c(2012, 1), end = c(2012, nTrain))
valid.ts <- window(gasoline.ts, start = c(2012, nTrain + 1))

gasolineprice.nnetar.opt <- nnetar(train.ts)

gasolineprice.nnetar.opt.pred <- forecast(gasolineprice.nnetar.opt, h = 9)

autoplot(train.ts, main = "Forecasts from Neural Network Method") +
  autolayer(valid.ts, series = "Validation") +
  autolayer(gasolineprice.nnetar.opt.pred, series = "Predictions")
accuracy(gasolineprice.nnetar.opt.pred, valid.ts)  
```

## Regression with external factors
```{r}
pairs(df1)
```

```{r}
qplot(df1$`CPI(ALL)`,df1$Gasoline_Retail_Price)
qplot(df1$`CPI(Gas)`,df1$Gasoline_Retail_Price)
qplot(df1$Crude_Oil_Price, df1$Gasoline_Retail_Price)
qplot(df1$Crude_Oil_Production,df1$Gasoline_Retail_Price)
```


## Study the cross-correlations between the each variable and gasoline retail price
```{r}
Ccf(df1$`CPI(ALL)`,df1$Gasoline_Retail_Price, 50)
Ccf(df1$`CPI(Gas)`,df1$Gasoline_Retail_Price, 77)
Ccf(df1$Crude_Oil_Price, df1$Gasoline_Retail_Price, 78)
Ccf(df1$Crude_Oil_Production,df1$Gasoline_Retail_Price, 50)
```

1. CPI(ALL) 34 months ago affect the gasoline retail price now

2. CPI(GAS) 76 months ago affect the gasoline retail price now

3. Spot price of crude oil contracts 77 months ago affect the gasoline retail price now

4. Crude oil production 20 months ago affect the gasoline retail price now

```{r}
library(tidyverse)
lag <- stats::lag

newdata <- ts.intersect(retail_price = gasoline.ts, 
                        leadcpi=lag(cpi.ts,-34),
                        leadcpigas=lag(cpi.gas.ts,-76),
                        leadcrudepri=lag(crudeoil.ts,-77),
                        leadcrudepro=lag(production.ts,-20))


m3 = tslm(retail_price~leadcpi + leadcpigas + leadcrudepri + leadcrudepro,data=newdata)
summary(m3)

m4 = tslm(retail_price~leadcpi + leadcrudepri + leadcrudepro,data=newdata)
summary(m4)
accuracy(m4$fitted.values, gasoline.ts)
```

```{r}
e <- m4$fitted.values - gasoline.ts
m <- quantile(e, probs = c(0.025,0.975))
lb=m4$fitted.values+m[1]
ub=m4$fitted.values+m[2]

reg.ts <- window(gasoline.ts,start=c(2018,6), frequency=12)
autoplot(m4$fitted.values, series = "Prediction", lwd=1.5, main="Forecasts from Multiple Regression") +
  autolayer(lb,lty="dashed")+ autolayer(ub,lty="dashed") + autolayer(reg.ts, series="Observed",lwd=0.5, col = 0)
```

```{r}
data = as.data.frame(newdata)
attach(data)
checkresiduals(m4)
acf2(resid(m4))

m5 = Arima(retail_price,order=c(2,0,0), xreg = cbind(leadcpi,leadcrudepri,leadcrudepro))
m5.pred = forecast(m5, xreg = cbind(leadcpi,leadcrudepri,leadcrudepro)) 
accuracy(m5.pred,retail_price)
```

## Compare all models
```{r}
accuracy(naive, validation)

accuracy(snaive, validation)

accuracy(ma.pred, validation)

accuracy(zzz.pred,validation)

accuracy(m1.pred, validation) #ARIMA(0,1,1)(1,0,0)12

accuracy(m2.pred, validation) #ARIMA(0,1,1)(0,0,0)12

accuracy(gasolineprice.nnetar.opt.pred, valid.ts)  

accuracy(m4$fitted.values, gasoline.ts)

AIC(m4)
m5$AIC
```

## Revaluate model accuracy with new released data
```{r}
actual.gasoline <- c(3.935,3.799)
actual.ts <- ts(actual.gasoline, start=c(2022,10),end=c(2022,11), frequency = 12)

#1
naive.pred = naive(training, h = 11)
accuracy(naive.pred, actual.ts)
accuracy(naive, validation)

#2
snaive.pred = snaive(training, h = 11)
accuracy(snaive.pred, actual.ts)
accuracy(snaive, validation)

#3
n = length(gasoline.ts)

w = 9
nValid = w

ma.trail.pred1= rep(NA, nValid)

for (i in 1:11) {
  nTraining= n-nValid + (i - 1)
  training3=window(gasoline.ts, start=c(2012, 1), 
                    end=c(2012, nTraining))
  
  ma.trailing.tmp1 = rollmean(training3, k=w, align="right")
  last.ma1=tail(ma.trailing.tmp1,1)
  
  ma.trail.pred1[i] = last.ma1
}

ma.pred1= ts(ma.trail.pred1,start=c(2012,length(gasoline.ts)-nValid+1), frequency = 12)

validation <- window(gasoline.ts, start=c(2022,1))

accuracy(ma.pred1, actual.ts)


#4
zzz.pred1=forecast(zzz,h=11,level=0)
accuracy(zzz.pred1, actual.ts)

#5
m1.pred1 = forecast(m1,h=11)
accuracy(m1.pred1, actual.ts)


#5
m2.pred1 = forecast(m2,h=11)
accuracy(m2.pred1, actual.ts)

#6
gasolineprice.nnetar.opt.pred1 <- forecast(gasolineprice.nnetar.opt, h = 11)
accuracy(gasolineprice.nnetar.opt.pred1, actual.ts)  


#7
summary(m4)
actual.gasoline <- c(3.935,3.799)


variable.data <- as.data.frame(ts.intersect(leadcpi1=lag(cpi.ts,-34),
                        leadcrudepri1=lag(crudeoil.ts,-77),
                        leadcrudepro1=lag(production.ts,-20)))

attach(variable.data)
reg.pred <- rep(NA, length(leadcpi1))
n = length(reg.pred)

for (i in 1:n) {
  reg.pred[i] = m4$coefficients[1] + m4$coefficients[2] * leadcpi1[i] +
                m4$coefficients[3]* leadcrudepri1[i] +
                m4$coefficients[4]* leadcrudepro1[i]
}
m4$fitted.values
reg.pred.ts <- ts(reg.pred, start=c(2018,6),frequency=12)
gasoline.ts <- ts(append(as.numeric(gasoline.ts), actual.gasoline),start=c(2012,1),frequency = 12)
accuracy(reg.pred.ts, gasoline.ts)
accuracy(reg.pred.ts, actual.ts)


e <- reg.pred.ts - gasoline.ts
m <- quantile(e, probs = c(0.025,0.975))
lb=reg.pred.ts+m[1]
ub=reg.pred.ts+m[2]

autoplot(reg.pred.ts) +
  autolayer(lb)+ autolayer(ub)

checkresiduals(m4)
qqnorm(e)
qqline(e)
```






